Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-04-15 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 15/04/25

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain in equilibrium 
while below a specific concentration ($C_{eq}$) - below 
which the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate the two 
phases. Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases. 
Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, and 
$C_{eq}(T)$ is the equilibrium concentration at temperature 
$T$. While the solution remains supersaturated there 
is a chemical potential driving the solute to coalesce 
and separate from the solution as an ordered solid, the 
first formation of the crystal is referred to as the 
nucleus and understanding its formation has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation 
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can 
be broadly categorised into either primary of secondary
nucleation. The former describes the formation of an initial
nucleus within the bulk phase, absent of any external stimuli.
Primary nucleation is therefore considered stochastic as there 
is no reliable means of predicting where a nucleus may form, 
or how long it will take. The only reliable indicator being 
that higher supersaturations will result in faster nucleation
rates. At a small scale one can estimate the nucleation rate 
by making repeated measurements of sample solutions and seeing
how many have nucleated after a given time, giving us a Poisson
probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
$V$ is the volume of the individual samples, and $M^*(t)$ \& 
$M$ are the number of nucleated samples and the total number 
of samples used respectively. While this is useful for studying
the effects of different parameters at a small scale, for
industrial applications there are to many external factors 
for primary nucleation to be measured accurately.

@@

\section{Nucleation Theories}
Several potential theories have been proposed to explain 
how nucleation occurs at a microscopic scale. In doing so 
we could potentially predict the expected yield of a given
crystalliser based on the initial conditions of the solution. 
Outlined below are some of the most popular theories currently 
used to describe nucleation.

@@

Consider a supersaturated solution, after some time 
enough individual sub units collide, forming a nucleus 
of volume $4\pi r^3/3$. The newly formed phase has a 
lower chemical potential than the surrounding solution, 
reducing the free energy of the system. Simultaneously, 
the formation of a new interface is resited by the bulk 
phase due to surface tension. The net free energy of the 
system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
	\label{eq:CNT} 
\end{align}

@@

CNT is often regarded as a good description of the macro system, its 
obvious that for all crystallization systems there is an inherent energy 
barrier that dictates the nucleation rate. Where it falters is in its 
predictive ability, both in estimating nucleation rates \cite{Gharibeh2005, 
Vekilov2010}, and in the structure of newly formed nuclei \cite{Lee1999, 
Yau2001}. Recent studies suggest classical nucleation is merely one of many 
possible pathways that can be taken to produce a structured crystalline phase. 
Prompting the development of alternative theories to better describe the 
nucleation process.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT 
that suggests that prior to nucleation, the solute will 
form precursor structures. The CNT assumes that solute 
structures below the critical radius should be unstable 
within the solution due to interfacial forces trying to 
dissolve the solute. This assumption is based on the idea 
that any interactions between solute molecules are much
smaller than the surface tension due to the bulk solution.
If instead the solute molecule interactions are significant
then it stands to reason that subcritical liquid-like 
structures are stable.

@@

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable fora single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice). There have 
been several different methods deployed to try and observe 
crystallisation at a microscopic scale.

@@

\subsection{Literature related to laser induced nucleation}
From as early as 1996 it has been known that laser irradiation 
using a Gaussian beam is a viable method of inducing nucleation 
within a supersaturated solution \cite{Garetz1996}. The first 
reported case was notable as it used a 1.064 $\mu m$ laser, the 
glycine solutions would appear transparent to such a laser which 
would suggest there was no photo-chemical reaction. Later studies 
into this phenomena found that the laser polarisation can influence 
the polymorph produced. With circularly polarised light producing 
$\alpha$-glycine and linearly polarised light forming $\gamma$-glycine 
\cite{Garetz2002}. Future research has found nucleation can be 
induced by 1 of 3 routes.

@@

\subsubsection{Trapping Induced Nucleation}
Lastly, there is trapping induced nucleation, this is where optical 
tweezers come into play. Due to the radiation pressure created by 
the focused beam, it is possible to manipulate the solute, this was 
demonstrated with amino acids such as glycine \cite{Tsuboi2009}. 
Whether or not a crystal forms is due to the location of the laser 
focus. When focusing on the cover slip, supersaturated solutions 
of glycine and $D_2O$ were shown to create a dense liquid droplet 
of glycine and water \cite{Yuyama2010, Yuyama2012}. The dipole 
moment of the glycine molecules is too small to be influenced by 
the optical trap, as such it would suggest that larger aggregates 
are being manipulated. Applying DLS analysis to the dense liquid 
region showed that it was populated by clusters that would 
consolidate together upon being focused by the optical trap 
\cite{Gowayed2021}. Molecular simulations of glycine solutions 
showed that these clusters are unstable when using pure glycine 
below the saturation point suggesting that the clusters are 
formed due to glycine reaction products \cite{Sweatman2022}. 
When the optical trap is moved from the cover slip to the 
air-solution interface, nucleation would occur before a dense 
liquid region could form \cite{Yuyama2010}. Repeated experiments 
where the laser is focused on the air-solution interface have 
lead to a variety of different nucleation events. In some 
instances the nucleation occurs spontaneously after a short 
period of time \cite{Yuyama2010}. Whereas allowing a solution to 
age results in the formation of amorphous precursors that when 
irradiated will nucleate immediately \cite{Liao2022}. The 
precursors are only seen when the solution is irradiated by an 
optical tweezer and the growth rate can be controlled somewhat 
by varying the laser power \cite{Liao2022}. The reason why 
nucleation is only seen at the air-solution interface is due 
to the limited molecular mobility close to the interface. 
Often tweezing experiments will use a hydrophilic coating to 
minimise the height of the solution droplet and further limit
the molecular mobility \cite{Yuyama2012, Gowayed2021}.

@@

Laser induced nucleation has the potential to be a viable method 
for \textit{in-situ} studying of nucleation events. Using high 
numerical aperture lens one can localise the nucleation event 
to a specific region of the solution. The current issue is that 
the mechanism behind laser induced nucleation is not fully 
understood, as such it is rather difficult to modify the laser 
for different solution parameters. Instead it may be more 
effective to manipulate the solution using trapped particles. 
One way would to generate an optical torque on a trapped particle
and therefore shear the surrounding fluid, a method that is 
already in common use for micro-rheological studies \cite{Bishop2004, 
RobertsonAnderson2018}

\subsection{Optical Torque and rotation}
\label{sec:opt_torque}
It has long been known that electromagnetic fields can transfer
linear and angular momentum \cite{Beth1936MechanicalDA}; more 
accurately the field is said to have both orbital and spin momentum. 
Though there is some debate on how to decompose the total momentum 
into these two components \cite{Bruce2020, Svak2018}, for this 
project we do not need to calculate the exact quantities and will 
instead look at the broader effects of both components. Orbital 
angular momentum arises from the shape of the wavefront of the 
particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation 
pressure that Ashkin and co demonstrated \cite{Ashkin1980}. However, 
higher order modes of a Gaussian beam (for example: Laguerre-Gaussian 
modes) have non-uniform wave fronts meaning the orbital momentum has 
both angular and linear components; depending on the relative size 
of the target particle one can induce rotation, or orbiting 
\cite{Bruce2020, Courtial2000}.

@@

\subsection{Characterisation of rotational motion}

Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres. The only drawback 
admitted to in their work was that the technique was constrained 
to isotropic scatters as their theoretical model for describing 
the auto-correlation function was predicated on the fact that 
any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical 
trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
However, as demonstrated by the results from Chapter 4.1, 
the axial traps of spherical aggregates is often situated 
far beyond the Rayleigh range (for a 1.2 NA laser this is 
$\pm5.985 \mu m$).

@@

\section{Significance of Thesis}
As I have hoped to make clear in the above introduction, the 
current state of nucleation theory is rather cumbersome at a 
micro-level. Models such as CNT and multi-step nucleation are 
not sufficient for describing the myriad of potential pathways 
nucleation can go down. As suggested by some review articles, 
the best way to address this is by developing \textit{in-situ} 
methods that can study the pre nucleation phase in greater detail 
\cite{Fu2021}. Furthermore, the ability to localise nucleation 
allows for better characterisation of the kinetics of crystal 
growth. Laser induced nucleation stands to be an ideal method to 
study nucleation, as the laser output can be concentrated to a 
small area \cite{Korede2023}. This fine control over the local 
fluid would allow for individual nucleation events to be
characterised as a factor of the local fluid properties. One 
challenge lies in the fact that its clear that the local fluid 
properties have a direct influence on the likelihood of laser
induced nucleation from occurring \cite{Korede2023, Ward2016,
Yuyama2012, Liao2022}.

>>>>

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. When 
a small amount of solute is added to the solvent the 
solute will diffuse throughout the solvent, reaching
an equilibrium concentration throughout the solution. 
Because of molecular interactions, the two can only remain 
in equilibrium while below a specific concentration ($C_{eq}$) 
- below which the chemical potential $\mu$ for a single phase 
is greater than the potential required to separate the two. 
Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases. 
Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

Where $C_{sol}$ is just the solute concentration, and 
$C_{eq}(T)$ is the equilibrium concentration of the solute  
at temperature $T$ based on the main solvent. While the 
solution remains supersaturated there is a chemical 
potential driving the solute to coalesce and separate 
from the solution as an ordered solid, the first 
formation of the crystal is referred to as the 
nucleus and understanding its formation has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation 
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can 
be broadly categorised into either primary of secondary
nucleation. The former describes the formation of an initial
nucleus within the bulk phase, absent of any external stimuli.
Primary nucleation is therefore considered stochastic as there 
is no means of predicting where a nucleus may form, or how long 
it will take. The only reliable indicator being that higher 
supersaturations will result in faster nucleation rates. At a 
small scale one can estimate the nucleation rate by making 
repeated measurements of sample solutions and seeing how many 
have nucleated after a given time, giving us a Poisson probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
	\label{eq:poisson}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
$V$ is the volume of the individual samples, and $M^*(t)$ \& 
$M$ are the number of nucleated samples and the total number 
of samples used respectively. The nucleation rate is one of 
the most important factors used in evaluating a crystallisers 
performance, providing a rough estimation of the final product 
mass over a given time frame. While \eqref{eq:poisson} is useful 
for studying the effects of different parameters at a small scale, 
for industrial applications there are to many external factors 
for primary nucleation to be measured accurately.

@@

\section{Nucleation Theories}
In order to predict the nucleation rate of a given solution
we need a solid understanding of how the individual molecules
interact such that they transition from an unordered phase to
a ordered phase. Over the years several theories have been 
proposed to explain how nucleation events occur, so far there
has yet to be a standout candidate that reliably predicts 
the systems behaviour. Either they fail to accurately predict
nucleation rates or more generally cannot describe new 
experimental results.

@@

Consider a supersaturated solution, at some time individual 
sub units collide rapidly, one after another, forming a 
nucleus of volume $4\pi r^3/3$. The newly formed phase has a 
lower chemical potential than the surrounding solution, 
reducing the free energy of the system. Simultaneously, 
the formation of a new interface is resited by the bulk 
phase due to surface tension. The net free energy of the 
system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
	\label{eq:CNT} 
\end{align}

@@

CNT is often regarded as a good description of the macro system, 
its obvious that for all crystallization systems there is an 
inherent energy barrier that dictates the nucleation rate. 
Where it falters is in its predictive ability, both in estimating 
nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
structure of newly formed nuclei \cite{Lee1999, Yau2001}.
Recent studies suggest classical nucleation is merely one of 
many possible pathways that can be taken to produce an ordered 
crystalline phase. Prompting the development of alternative 
theories to better describe the nucleation process.

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT 
that suggests that prior to nucleation, the solute will 
form precursor structures. The CNT assumes that solute 
structures below the critical radius should be unstable 
within the solution due to interfacial forces trying to 
dissolve the solute. This assumption is based on the idea 
that any interactions between solute molecules are much
weaker than that between solute and solvent molecules.
If instead the solute molecule interactions are significant
then it stands to reason that subcritical liquid-like 
structures are stable.

@@

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable for a single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice). There have 
been several different methods deployed to try and observe 
crystallisation at a microscopic scale.

@@

\subsection{Laser induced nucleation}
From as early as 1996 it has been known that laser irradiation 
using a Gaussian beam is a viable method of inducing nucleation 
within a supersaturated solution \cite{Garetz1996}. The first 
reported case was notable as it used a 1.064 $\mu m$ laser, the 
glycine solutions would appear transparent to such a laser which 
would suggest there was no photo-chemical reaction. Later studies 
into this phenomena found that the laser polarisation can influence 
the polymorph produced. With circularly polarised light producing 
$\alpha$-glycine and linearly polarised light forming $\gamma$-glycine 
\cite{Garetz2002}. Future research has found nucleation can be 
induced by 1 of 3 routes.

@@

\subsubsection{Trapping Induced Nucleation}
Lastly, there is trapping induced nucleation, this is where optical 
tweezers come into play (see below). Due to the radiation pressure 
created by the focused beam, it is possible to manipulate the solute, 
this was demonstrated with amino acids such as glycine \cite{Tsuboi2009}. 
Whether or not a crystal forms is due to the location of the laser 
focus. When focusing on the cover slip, supersaturated solutions 
of glycine and $D_2O$ were shown to create a dense liquid droplet 
of glycine and water \cite{Yuyama2010, Yuyama2012}. The dipole 
moment of the glycine molecules is too small to be influenced by 
the optical trap, as such it would suggest that larger aggregates 
are being manipulated. Applying dynamic light scattering analysis to 
the dense liquid region showed that it was populated by clusters that 
would consolidate together upon being focused by the optical trap 
\cite{Gowayed2021}. Molecular simulations of glycine solutions 
showed that these clusters are unstable when using pure glycine 
below the saturation point suggesting that the clusters are 
formed due to glycine reaction products \cite{Sweatman2022}. 
When the optical trap is moved from the cover slip to the 
air-solution interface, nucleation would occur before a dense 
liquid region could form \cite{Yuyama2010}. Repeated experiments 
where the laser is focused on the air-solution interface have 
lead to a variety of different nucleation events. In some 
instances the nucleation occurs spontaneously after a short 
period of time \cite{Yuyama2010}. Whereas allowing a solution to 
age results in the formation of amorphous precursors that when 
irradiated will nucleate immediately \cite{Liao2022}. The 
precursors are only seen when the solution is irradiated by an 
optical tweezer and the growth rate can be controlled somewhat 
by varying the laser power \cite{Liao2022}. The reason why 
nucleation is only seen at the air-solution interface is due 
to the limited molecular mobility close to the interface. 
Often tweezing experiments will use a hydrophilic coating to 
minimise the height of the solution droplet and further limit
the molecular mobility \cite{Yuyama2012, Gowayed2021}.

@@

Laser induced nucleation has the potential to be a viable method 
for \textit{in-situ} studying of nucleation events. Using high 
numerical aperture lens one can localise the nucleation event 
to a specific region of the solution. The current issue is that 
the mechanism behind laser induced nucleation is not fully 
understood, as such it is rather difficult to modify the laser 
for different solution parameters. Instead it may be more 
effective to manipulate the solution using trapped particles. 
One way would to generate an optical torque on a trapped particle
and therefore shear the surrounding fluid, a method that is 
already in common use for micro-rheological studies \cite{Bishop2004, 
RobertsonAnderson2018}.

\subsection{Optical Torque and rotation}
\label{sec:opt_torque}
Electromagnetic fields can transfer both linear and angular momentum \cite{Beth1936MechanicalDA}; more accurately the field is said to have 
both orbital and spin momentum. Though there is some debate on how to decompose the total momentum into these two components \cite{Bruce2020, Svak2018}, for this project we do not need to calculate the exact 
quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the 
wavefront of the particular field in question; for simple Gaussian 
beams the wavefronts are uniform and equally spaced resulting in the 
typical radiation pressure that Ashkin and co demonstrated 
\cite{Ashkin1980}. However, higher order modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have non-uniform wave fronts meaning 
the orbital momentum has both angular and linear components; depending 
on the relative size of the target particle one can induce rotation, 
or orbiting \cite{Bruce2020, Courtial2000}.

@@

\subsection{Characterisation of rotational motion}
Rotational motion about a single axis is easiest to account for.  
When the power spectra of elliptical polystyrene particles was 
fitted by Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, they 
assumed that the rotational motion was purely in the transverse
plane. As such they did not have to account for any variance in 
the trapping strength due to orientation nor did they need to
consider non-periodic rotational behaviour. In the case where 
rotational motion is stochastic the problem is more complex. 
For example, when an optical fibre trap characterisation 
technique was implemented by Saffron \textit{et al} 
\cite{BarZiv1997, Meller1998}, they were able to use dynamic 
light scattering to characterise both the axial and lateral 
trap stiffness acting on microspheres. The only drawback 
admitted to in their work was that the technique was constrained 
to isotropic scatters as their theoretical model for describing 
the auto-correlation function was predicated on the fact that 
any variations in the signal are due to the particles 
translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. Where the upper limit of the cylindrical 
trap is given by the Rayleigh range ($z_R = n\omega_0/NA$).
However, as demonstrated by the results from Chapter 4.1, 
the axial traps of spherical aggregates is often situated 
far beyond the Rayleigh range (for a 1.2 NA laser this is 
$\pm5.985 \mu m$).

@@

\section{Significance of Thesis}
As I have hoped to make clear in the above introduction, the 
current state of nucleation theory is rather cumbersome at a 
micro-level. Models such as CNT and multi-step nucleation are 
not sufficient for describing the myriad of potential pathways 
nucleation can go down. As suggested by some review articles, 
the best way to address this is by developing \textit{in-situ} 
methods that can study the pre nucleation phase in greater detail 
\cite{Fu2021}. Furthermore, the ability to localise nucleation 
allows for better characterisation of the kinetics of crystal 
growth. Laser induced nucleation stands to be an ideal method to 
study the nucleation of organic compounds, as the laser output 
can be concentrated to a small area \cite{Korede2023} while not 
altering the organic compound - as is the case with TEM. However,
laser induced nucleation in itself is also poorly understood, 
meaning trying to study the nucleation kinetics is difficult if the
effect of the laser. With this in mind we suggest that we instead 
utilise another application of optical tweezers, manipulation 
of micro-scale particles to induce nucleation around a small 
localised area. The most common method for manipulating a fluid is 
by localised fluid shearing, which has been shown to directly 
influence nucleation rates \cite{Debuysschere2023}. 
One challenge lies in the fact that its clear that the local fluid 
properties have a direct influence on the likelihood of laser
induced nucleation from occurring \cite{Korede2023, Ward2016,
Yuyama2012, Liao2022}.
